---
title: "Analysis of Trend for NHANES Lab Test "
author: "Shu Deng"
date: "4/20/2020"
output:
  html_document:
    code_folding: hide
    css: "trend_analysis_2.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(data.table)
library(survey)
library(tidyverse)
library(plotly)
library(ggpubr) # for arrange plots
library(broom)
library(patchwork)
library(ggthemes)
theme_set(theme_tufte())
```

Data Scource: NHANES    
Pre process : See Git   

```{r}

survey_df = read_csv("data/nhanes_99_11_raw.csv")
survey_df$year_num = as.numeric(as.factor(survey_df$year) ) 
survey_df$year_num_2 = (as.numeric(survey_df$year )- 1999)/2
survey_df$year_fct = as.factor(survey_df$year) 
# survey_df[,.N]
# survey_df[,.N, by = year]
# plot(survey_df[,.N, by = year])
```

***

Design survey objects using PSU, Strata and Weights.    
Since we are not combining multiple years together, we use 2-Year medical examination weight.     
However, we do have to subset design for subpopulation: Children   

```{r}
# for all ages 
nhanes_svy = 
  svydesign(
    id =  ~ SDMVPSU,
    strata =  ~ SDMVSTRA,
    weights =  ~ WTMEC2YR,
    data = survey_df,
    nest = TRUE
  )

# for children 
nhanes_svy_kid = 
  subset(nhanes_svy,
         RIDAGEEX <= 18*12 & RIDAGEEX >= 12*12
    )

```



#### Compare Survey vs non-Survey weight 
<br/>
<br/>


<div>

<div class="container123">
<div class="fixed123">
    
    
For All Ages
```{r}
# weighted bar plot 
svyby_adult_1  =
  svyby(
    formula = ~ LBDLDL,
    by = ~ year_fct,
    design = nhanes_svy,
    FUN = svymean,
    na.rm = TRUE,
    keep.names = FALSE
  )
             
ggplot(
  data = svyby_adult_1, 
  aes(x = year_fct, y = LBDLDL)
  ) +
  geom_col() +
  geom_errorbar(
    aes(ymin = LBDLDL - 2 * se, ymax = LBDLDL + 2 * se),
    width = .2,
    position = position_dodge(.9)
  ) +
  labs(y = "Average Yearly LDL", x = "", title = "Weighted mean")

### unweighted all age 
survey_df %>% 
  group_by(year_fct) %>% 
  summarise(mean_ldl = mean(LBDLDL, na.rm = TRUE)) %>% 
  ggplot(
    aes(x = year_fct, y = mean_ldl)
    ) +
    geom_col() +
  labs(y = "Average Yearly LDL", title = "Unweighted/ Sample mean")

```

</div>
<div class="flex-item123">


For Children
```{r}
##  --- -- --  weighted Children  -- -- -  -- -## 
svyby_kid_1  =
  svyby(
    formula = ~ LBDLDL,
    by = ~ year_fct,
    design = nhanes_svy_kid,
    FUN = svymean,
    na.rm = TRUE,
    keep.names = FALSE
  )
             
ggplot(
  data = svyby_kid_1, 
  aes(x = year_fct, y = LBDLDL)
  ) +
  geom_col() +
  geom_errorbar(
    aes(ymin = LBDLDL - 2 * se, ymax = LBDLDL + 2 * se),
    width = .2,
    position = position_dodge(.9)
  ) +
  labs(y = "Average Yearly LDL", x = "", title = "Weighted mean")


### unweighted Children 

survey_df %>% 
  filter(RIDAGEEX <= 18*12 & RIDAGEEX >= 12*12) %>% 
  group_by(year_fct) %>% 
  summarise(mean_ldl = mean(LBDLDL, na.rm = TRUE)) %>% 
  ggplot(
    aes(x = year_fct, y = mean_ldl)
    ) +
    geom_col() +
  labs(y = "Average Yearly LDL", title = "Unweighted/ Sample mean")
```

</div>
</div>
</div>



```{r eval=FALSE, include=FALSE}
# ANOVA & pariwise comparison without using sampling weight
# build model 
ano_res_1 = aov(LBDLDL ~ year_fct , data = survey_df[RIDAGEYR %between% c(12, 18), ])
summary(ano_res_1)
TukeyHSD(ano_res_1)
# plot(ano_res_1)
```

#### For Survey Data
  
Year of the survey entered as a continuous variable, from 0 for NHANES 1999/2000 to 6 for NHANES 2011/2012.   

```{r fig.width=10}
box_plot_1 =
  ggplot(data = survey_df,
         mapping = aes(x = year_num_2, y = LBDLDL)) +
  geom_boxplot(aes(group = year_num_2)) +
  guides(color = FALSE) +
  geom_smooth(method = "lm",
              se = TRUE,
              mapping = aes(weight = WTMEC2YR)) +
  labs(title = "For All ages ")


box_plot_2 =
  ggplot(data = filter(survey_df, RIDAGEEX <= 18*12 & RIDAGEEX >= 12*12),
       mapping = aes(x = year_num_2, y = LBDLDL)) +
  geom_boxplot(aes(group = year_num_2)) +
  guides(color = FALSE) +
  geom_smooth(method = "lm",
              se = TRUE,
              mapping = aes(weight = WTMEC2YR)) +
  labs(title = "For Children ")

box_plot_1 + box_plot_2
```



Compare weighted mean of LDL. Linear & Smooth    
```{r fig.width=10}
point_plot_1 = 
  svyby( ~ LBDLDL , ~ year, nhanes_svy_kid, svymean, na.rm = TRUE) %>%
  ggplot(aes(x = year, y = LBDLDL)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = LBDLDL - 2 * se, ymax = LBDLDL + 2 * se),
    width = .2,
    position = position_dodge(.9)
  )  +
  geom_smooth(
    data = filter(survey_df, RIDAGEEX <= 18*12 & RIDAGEEX >= 12*12),
    method = "lm",
    se = TRUE,
    mapping = aes(x = year , y = LBDLDL, weight = WTMEC2YR)
  ) +
  labs(y = "Average Yearly LDL")



point_plot_2 = 
svyby(~ LBDLDL , ~ year, nhanes_svy_kid, svymean, na.rm = TRUE) %>%
  ggplot(aes(x = year, y = LBDLDL)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = LBDLDL - 2 * se, ymax = LBDLDL + 2 * se),
    width = .2,
    position = position_dodge(.9)
  )  +
  geom_smooth(
    data = filter(survey_df, RIDAGEEX <= 18 * 12 & RIDAGEEX >= 12 * 12),
    method = "loess",
    se = TRUE,
    mapping = aes(x = year , y = LBDLDL, weight = WTMEC2YR)
  ) +
  labs(y = "Average Yearly LDL")

point_plot_1 + point_plot_2
```


```{r include=FALSE}
# local polynomial smoothing 
smth <- svysmooth(LBDLDL ~ year_num ,nhanes_svy_kid,bandwidth= 1 , method = "locpoly")
plot(smth)
```

For four types of lipid analytes
```{r}
yearly_avg_kid_ldl = svyby(~ LBDLDL , ~ year, nhanes_svy_kid, svymean, na.rm = TRUE)
yearly_avg_kid_hdl = svyby(~ LBDHDD , ~ year, nhanes_svy_kid, svymean, na.rm = TRUE)
yearly_avg_kid_tc = svyby(~ LBXTC , ~ year, nhanes_svy_kid, svymean, na.rm = TRUE)
yearly_avg_kid_tg = svyby(~ LBXTR , ~ year, nhanes_svy_kid, svymean, na.rm = TRUE)

yearly_avg_kid_ldl$type = "LDL"
colnames(yearly_avg_kid_ldl)[2] = "value"

yearly_avg_kid_hdl$type = "HDL"
colnames(yearly_avg_kid_hdl)[2] = "value"

yearly_avg_kid_tc$type = "TC"
colnames(yearly_avg_kid_tc)[2] = "value"

yearly_avg_kid_tg$type = "TG"
colnames(yearly_avg_kid_tg)[2] = "value"

yearly_avg_kid_combo = bind_rows(list(yearly_avg_kid_ldl,yearly_avg_kid_hdl,yearly_avg_kid_tc,yearly_avg_kid_tg))
yearly_avg_kid_combo$year = as.numeric(yearly_avg_kid_combo$year)


yearly_avg_kid_combo %>% 
  ggplot() + 
  geom_line(aes(x = year , y = value , color = type)) +
  geom_errorbar(aes(x = year, ymin = value - 2 * se, ymax = value + 2 * se, width = .1, color = type),position = position_dodge(.1)) +
  labs(x = "" , y = "Average Yearly Cholesterols Level", title = "For Children")
```


```{r}
# 2 X 2 bar plot
ldl_bar_plot = 
  yearly_avg_kid_ldl %>% 
  ggplot(aes(x = year, y = value)) +
  geom_col() + 
  geom_errorbar(aes(ymin = value - 2 * se, ymax = value + 2 * se, width = .1)) +
  labs(x = "", y = "")

tc_bar_plot = 
  yearly_avg_kid_tc %>% 
  ggplot(aes(x = year, y = value)) +
  geom_col() + 
  geom_errorbar(aes(ymin = value - 2 * se, ymax = value + 2 * se, width = .1))+
  labs(x = "", y = "")

hdl_bar_plot = 
  yearly_avg_kid_hdl %>% 
  ggplot(aes(x = year, y = value)) +
  geom_col() + 
  geom_errorbar(aes(ymin = value - 2 * se, ymax = value + 2 * se, width = .1))+
  labs(x = "", y = "")

tg_bar_plot = 
  yearly_avg_kid_tg %>% 
  ggplot(aes(x = year, y = value)) +
  geom_col() + 
  geom_errorbar(aes(ymin = value - 2 * se, ymax = value + 2 * se, width = .1))+
  labs(x = "", y = "")





figure <- ggarrange(tc_bar_plot, ldl_bar_plot, hdl_bar_plot, tg_bar_plot, labels = c("TC", "LDL", "HDL","TG"))
annotate_figure(figure,
                # top = text_grob("For Children", color = "black", face = "bold", size = 12),
                # bottom = text_grob("Data source: \n ToothGrowth data set", color = "black",
                #                    hjust = 1, x = 1, face = "italic", size = 10),
                left = text_grob("Yearly Average Level (mg/dL) For Children", color = "black", rot = 90),
                # right = "I'm done, thanks :-)!",
                fig.lab = "", fig.lab.face = "bold"
)


```


```{r eval=FALSE, include=FALSE}
# check for normality 
survey_df$LBXTR %>% hist() 
```


```{r}
# Trend test for all analytes 
md_ldl <- svyglm(LBDLDL ~ year_num_2 + as.factor(RIDRETH1) + as.factor(RIAGENDR) + RIDAGEEX, design = nhanes_svy_kid)
md_hdl <- svyglm(LBDHDD ~ year_num_2 +as.factor( RIDRETH1) + as.factor(RIAGENDR) + RIDAGEEX, design = nhanes_svy_kid)
md_tg  <- svyglm(log(LBXTR) ~ year_num_2 + as.factor(RIDRETH1) + as.factor(RIAGENDR) + RIDAGEEX, design = nhanes_svy_kid)
md_tc  <- svyglm(LBXTC ~ year_num_2 + as.factor(RIDRETH1) + as.factor(RIAGENDR) + RIDAGEEX, design = nhanes_svy_kid)
```

#### Linear trend analysis result 
```{r}
trend_res = 
  bind_rows(
  list(LDL = tidy(md_ldl)[2, ], 
       "HDL" = tidy(md_hdl)[2, ], 
       TC  = tidy(md_tc)[2, ], 
       "*TG"  = tidy(md_tg)[2, ]),.id = "analyte") 
 knitr::kable(trend_res[,-2])
 
```

\*TG is log transformed due to non gaussian distritbution

